Blaming the Typeless: \\ Scalable, Human-Centric Python Localization

### 2018-07-13 paper/main.tex
add pipeline outline

>>>>

\emph{(Data pipeline summary)} Our raw data consist of every Python 3 program
that a user executed on PythonTutor.com (not in "live" mode) during 2017, which
comes to around 5 million programs. Each program which crashes is paired with
the next program (by the same user) that does not crash. We discard pairs where the
difference between crashing and fixed versions is too high (more than a stddev
above average).

For each statement and expression in each
program, we compute features including the following:

\begin{itemize}

\item \emph{Static / syntactic features}
\begin{enumerate}
    \item What kind of statement/expression it is, e.g. assignment / return
     / import (for statements) or variable / literal / application (expressions).
    \item Size (the number of AST nodes in the subtree rooted at this node)
\end{enumerate}

\item \emph{Dynamic features}
We run each program through an instrumented interpreter (i.e. a modified PythonTutor
backend, which itself is based on the BDB debugging library). This lets us
compute the following features:
\begin{enumerate}
    \item What type(s) does the expression have at runtime?
    \item Is it part of the error slice?
\end{enumerate}
(Since this interpreter works at the granularity of lines and we want to work with
expressions as well, we first convert each program to ANF.)

We compute the error slice by running the program and building a dependency
graph where node A depends on node B if either B defines a variable that A uses,
or B affects control flow allowing A to run.

We also check what exception is thrown (TypeError, KeyError, etc), and add that
as a feature to all vectors derived from that program.

\item \emph{Contextual features} After all other features are computed so we have a
vector $v_e$ representing expression $e$, we set $v_e^{+context} = v_e \circ
v_p \circ v_{c_1} \circ v_{c_2} \circ v_{c_3}$ where $p$ is the expression's parent
in the AST and $c_i$ are its children. This allows us to recover some of the
program structure that would otherwise be completely lost when we convert the
structured program into an unstructured bag of feature vectors.

\item \emph{Label} Each feature vector is labeled by whether that expression/statement
changed between the crashing and fixed versions. (This is our proxy for whether
that node is to blame for the crash).

\end{itemize}

Each vector is then one-hot encoded (e.g. feature "Type : Int" becomes set
of features "Type-Int : 1", "Type-Bool : 0", "Type-Str : 0", etc.) for use with
standard machine learning tools. The models we train (inc. decision trees/forests,
MLP, GBM) can then rank expressions of a test program in order of how likely they
are to be the source of the bug, so we can compute Top-k accuracies: how often
is the true program change located in the model's first k outputs. We report Top-1
and Top-3...
